Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2020-01-02 manuscript/abstract.tex
Edit abstract to 150 words or less
This is 149 words according to emacs word count and 144 according to wc.

\section*{Key findings}

\begin{itemize}
\item{Our model accurately forecasts clade frequency trajectories and identifies optimal potential vaccine strains for both simulated and natural populations by estimating the sequence composition of future populations and without relying on clade-based model targets}
\item{Experimental measurements of antigenic drift based on HI assays are more robust and predictive of viral success than epitope mutations}
\item{Models based on a predefined set of epitope sites risk overfitting to historical data}
\item{The combination of non-epitope mutation and LBI metrics outperforms all experimentally-informed metrics on validation data (Oct 1990--Oct 2015) but fails to account for reassortment events in test data (Oct 2015--Oct 2019)}
\item{The combination of HI assays and non-epitope mutations performs best on test data and is among the best models for validation data}
\item{Deleterious mutations contribute more to seasonal influenza evolution than has been as widely appreciated}
\item{DMS measurements appear to be too background-specific to be predictive for global viral success}
\item{Our model live forecasts in nextstrain.org will aid in year-round surveillance of influenza evolutionary patterns and allow us to continuously evaluate model performance relative to recent observations}
\item{Our model is the first of its kind to be released as an open source framework that can be inspected and extended by others}
\item{Immediate next steps to improve influenza models under this framework include the integration of geographic information and antigenic escape assay data}
\item{Further efforts to understand the declining efficacy of HI assays and to replace these with FRA or antigenic escape assays could improve models in the future}
\item{Our distance-based model targets and easy definition of fitness metrics through tidy data frames paves the way for future forecasting efforts with pathogens that cannot be analyzed by standard phylogenetic methods (e.g., highly recombinant viruses and organisms with larger genomes like bacteria and fungi)}
\end{itemize}

>>>>

Seasonal influenza virus A/H3N2 is a major cause of death globally.
Vaccination remains the most effective preventative.
Rapid mutation of hemagglutinin allows viruses to escape adaptive immunity.
This antigenic drift necessitates regular vaccine updates.
Effective vaccine strains need to match H3N2 populations circulating one year after strain selection.
Experts select strains based on experimental measurements of antigenic drift and predictions made by models from hemagglutinin sequences.
No modern models use phenotypic measures of viral fitness in their predictions.
We developed a novel influenza forecasting framework that integrates phenotypic measures of antigenic drift and functional constraint with previously published sequence-only fitness estimates.
Forecasts informed by phenotypic measures of antigenic drift consistently outperformed previous sequence-only estimates, while sequence-only estimates of functional constraint surpassed more comprehensive experimentally-informed estimates.
Importantly, the best models integrated estimates of both functional constraint and either antigenic drift phenotypes or recent population growth.
